### 2018-11-05 1/exams/m2.tex
Midterm 2.

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2017 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 15 \\ \hline
  2 & 12 \\ \hline
  3 & 20 \\ \hline
  4 & 20 \\ \hline
  5 & 15 \\ \hline
  6 & 20 \\ \hline\hline
  Total: & 102\\
  \hline
\end{tabular}
\end{center}

@@

\noindent
{\bf Problem 1 (15 points).}
%
Here are some data definition relevant to representing a dictionary,
which associates words with the their definitions:

\begin{verbatim}
;; A Dict is one of:
;; - '()
;; - (cons (list String String) Dict)
;; Interp: represents a collection of definitions where each
;; element is a two-element list consisting of a word (first) and its
;; meaning (second).

;; A Result is one of:
;; - #false
;; - String
;; Interp: represents the result of looking up a word in the
;; dictionary; #false means not found.
\end{verbatim}

\noindent
Design the following function:

\begin{verbatim}
;; lookup : String Dict -> Result
;; Lookup the (first) definition of a word in the dictionary,
;; produces #false if not found
\end{verbatim}

@@

(define (f b s) s)
(define (g fs) (fs #false))
(define (h bs ns) 2)
(define (i f) (f 2))
(define (j f xs) (map f xs))
(define (k f g x) (g (f x)))
\end{SaveVerbatim}

%\censor{\vspace{-2.5in}}
\censor{%
\noindent\BUseVerbatim{VerbEnv}}

\noindent
{\bf Problem 2 (12 points).}
%
For each of the following functions, provide the most general signature
that correctly describes the function:

\begin{alltt}
;; method : \censor{Number -> Number}

(define (method x)
  (+ (sqr x) 2))

;; raekwon : \censor{[Listof Number] -> Boolean}

(define (raekwon x)
  (ormap positive? x))

;; rza : \censor{[Listof String] [Listof Number] -> [Listof Number]}

(define (rza x y)
  (cond [(empty? x) y]
        [(cons? x)
         (cons (string-length (first x))
               (rza (rest x) y))]))

;; gza : \censor{(String Number -> Number) -> Number}

(define (gza x)
  (foldr x 0 (list "a" "b" "c")))

;; odb : \censor{(Number -> Number) -> (Number -> Number)}

(define (odb x)
  (lambda (y)   
    (/ (- (x (+ y 0.001))
          (x (- y 0.001)))
       (* 2 0.001))))

;; ghostface : \censor{[X Y] [Listof X] [X -> Y] -> [Listof Y]}

(define (ghostface x y)
  (cond [(empty? x) '()]
        [(cons? x)
         (cons (y (first x)) 
               (ghostface (rest x) y))]))
\end{alltt}

\noindent
{\bf Problem 3 (20 points).}
%
Design a program called {\tt w-avg} that computes the weighted average
of a list of numbers and a list of weights.

For example, let's say a class grade is based on two midterms and a
project where the project is worth twice as much as the midterms.  A
student who gets an 80 and a 70 on the midterms and a 90 on the
project would have a weighted average of 82.5: $((1 \times 80) + (1
\times 70) + (2 \times 90)) / 4$, which can be computed with {\tt
  (w-avg (list 80 70 90) (list 1 1 2))}.  You should assume the two
lists have the same length.

;; avg : [Listof Number] -> Number
;; Compute the average of a list of numbers
;; Assume: the list is non-empty
(check-expect (avg (list 3 4 5)) 4)
(check-expect (avg (list 10 20)) 15)
(define (avg lon)
  (/ (sum lon) 
     (length lon)))

;; sum : [Listof Number] -> Number
;; Sum the list of numbers
(check-expect (sum '()) 0)
(check-expect (sum (list 3 4 5)) 12)
(define (sum lon)
  (foldr + 0 lon))

;; Alt: follow [Listof Number] template
\end{SaveVerbatim}

\newpage 
\noindent
{\bf Problem 4 (20 points).}
%
Here is a parametric data definition for a tree of elements:
\begin{verbatim}
;; A [Tree X] is one of:
;; - (make-leaf)
;; - (make-node X [Tree X] [Tree X])
;; Interp: a binary tree that is either empty (a leaf), or non-empty (a node)
;; with an element and two sub-trees.
(define-struct leaf ())
(define-struct node (elem left right))
\end{verbatim}
{\bf Problem 4(a) (10 points).} Design a function, {\tt tree-sum},
that operates on a binary tree of numbers, {\tt [Tree Number]}, and
produces the sum of all the numbers in the tree.

;; tree-sum : [Tree Number] -> Number
;; Sum all the numbers in the tree
(check-expect (tree-sum (make-leaf)) 0)
(check-expect (tree-sum (make-node 7 (make-node 1 (make-leaf) (make-leaf)) 
                                     (make-leaf)))
              8)
(define (tree-sum bt)
  (cond [(leaf? bt) 0]
        [(node? bt)
         (+ (node-elem bt)
            (tree-sum (node-left bt))
            (tree-sum (node-right bt)))]))

@@

\noindent
{\bf Problem 4(b) (10 points).}
Here is a function that produces a list of all the elements in a tree:
\begin{verbatim}
;; tree-elems : [X] . [Tree X] -> [Listof X]
;; Produce a list of all elements in tree going top-down, left-to-right
(check-expect (tree-elems (make-leaf)) '())
(check-expect (tree-elems (make-node 7 
                                     (make-node 9 (make-leaf) (make-leaf))
                                     (make-node 2 (make-leaf) (make-leaf))))
              (list 7 9 2))
(define (tree-elems bt)
  (cond [(leaf? bt) '()]
        [(node? bt)
         (cons (node-elem bt)
               (append (tree-elems (node-left bt))
                       (tree-elems (node-right bt))))]))
\end{verbatim}
Here is an abstraction function for trees that is similar to
{\tt foldr} for lists, but works on trees:
\begin{verbatim}
;; tree-fold : [X Y] . [X Y Y -> Y] Y [Tree X] -> Y
;; The fundamental abstraction function for trees
(define (tree-fold f b bt)
  (cond [(leaf? bt) b]
        [(node? bt)
         (f (node-elem bt)
            (tree-fold f b (node-left bt))
            (tree-fold f b (node-right bt)))]))
\end{verbatim}
Give an equivalent definition of {\tt tree-elems} in terms of {\tt
  tree-fold}.  (Just provide the code.)

@@

(define (tree-elems bt)
  (tree-fold (lambda (x l r) (cons x (append l r))) '() bt))

@@

\newpage
\noindent
{\bf Problem 5 (15 points).}  
%
Design a program that takes a list of strings and produces the count
of strings with length longer than 3 in the list.  For example, if the
list contains {\tt "a"}, {\tt "dave"}, {\tt "abc"}, and {\tt "fred"},
the count is 2.  You may assume the {\tt{[Listof String]}} data
definition is defined.
%
For full credit, use list abstraction functions.  For partial credit,
follow the template for {\tt{[Listof String]}}.

@@

;; count-3+ : [Listof String] -> Number
;; Count the number of strings with length more than 3 in given list
(check-expect (count-3+ '()) 0)
(check-expect (count-3+ (list "a" "abcd" "fred" "c")) 2)
(define (count-3+ los)
  (foldr (lambda (s c) (if (> (string-length s) 3) (add1 c) c)) 0 los))

;; Alts:
(define (count-3+ los)
  (local [;; count+ : String Number -> Number
          ;; Bump count if string has length > 3
          (define (count+ s c) 
            (if (> (string-length s) 3) (add1 c) c))]
    (foldr count+ 0 los))

(define (count-3+ los)
  (local [;; 3+? : String -> Boolean
          ;; Does the string have length > 3?
          (define (3+? s) 
            (> (string-length s) 3))]
    (length (filter 3+? los))))

(define (count-3+ los)
  (length (filter (lambda (s) (> (string-length s) 3)) los)))
\end{SaveVerbatim}

@@

\noindent
{\bf Problem 6 (20 points).}
%
Here's a data definition for choose-your-own-adventure books:
\begin{verbatim}
;; A CYOA is one:
;; - String
;; - (make-choice String CYOA CYOA)
(define-struct choice (q yes no))
;; Interp: a choose-your-own-adventure book where a string
;; is a conclusion and a choice is a yes/no question and
;; the two adventures that follow as a consequnce of the answer.
\end{verbatim}

\noindent
Design a function called {\tt ans} that consumes a {\tt CYOA} and a
{\tt [Listof Boolean]} that represents answers to the yes/no
questions: {\tt \#true} means yes, {\tt \#false} means no.  The {\tt
  ans} function produces the {\tt CYOA} adventure that results after
giving all of the answers in the list.  If the list contains too many
answers, {\tt ans} just produces the conclusion of the adventure.

@@

(define a0 "A0")
(define a1 (make-choice "Q1" "A1" "A2"))
(define a2 (make-choice "Q2" a0 a1))

;; ans : CYOA [Listof Boolean] -> CYOA
;; ...
(check-expect (adv a0 (list #true #false)) "A0")
(check-expect (adv a1 (list #true #false)) "A1")
(check-expect (adv a2 (list #false #true)) "A1")
(check-expect (adv a2 (list #false)) a1)
(define (ans adv loa)
  (cond [(string? adv) adv]
        [(empty? loa) adv]
        [(first loa)
         (ans (choice-yes adv) (rest adv))]
        [else
         (ans (choice-no adv) (rest adv))]))
\end{SaveVerbatim}

@@

>>>>

\begin{center}
  {\Large {\bf CMSC 131A, Midterm 2}\\
    \censor{{\bf SOLUTION}} \ \\
    \ \\
    Fall 2018 \ \\
  }
\end{center}

@@

\begin{center}
\begin{tabular}{| c | c |}
  \hline
  Question & Points \\ \hline \hline
  1 & 10 \\ \hline
  2 & 15 \\ \hline
  3 & 15 \\ \hline
  4 & 20 \\ \hline
  5 & 20 \\ \hline\hline
  Total: & 80\\
  \hline
\end{tabular}
\end{center}

@@

\noindent
{\bf Problem 1 (10 points).}
%
Here is a data definition for representing definitions in a
dictionary:
\begin{verbatim}
;; A Defn is a (make-defn String String)
(define-struct defn (word meaning))
;; Interp: a word and its definition
;; Example:
(make-defn "pair" 
           "a set of two things used together or regarded as a unit")
\end{verbatim}

Design the following function for determining whether one definition
should come before another in a dictionary.  (Hint: you can use the
function {\tt string<?} to help.)
\begin{verbatim}
;; defn<? : Defn Defn -> Boolean
;; Does the word defined in d1 come before d2, alphabetically?
(define (defn<? d1 d2) ...)
\end{verbatim}

@@

(check-expect (defn<? (make-defn "a" "first") (make-defn "b" "second")) #true)
(check-expect (defn<? (make-defn "b" "second") (make-defn "a" "first")) #false)
(define (defn<? d1 d2)
  (string<? (defn-word d1) (defn-word d2)))

@@

\noindent
{\bf Problem 2 (15 points).}
%
A dictionary can be represented using a structure similar to a binary
search tree as follows (where {\tt Defn} is defined in problem 1):
\begin{verbatim}
;; A BSD (binary search dictionary) is one of:
;; - (make-leaf)
;; - (make-node d l r), where d is a Defn, l is a BSD, r is a BSD,
;;      and every definition in l alphabetically comes before d
;;      and d comes alphabetically before every definition in r
;; ASSUME: words are defined at most once in a BSD
(define-struct leaf ())
(define-struct node (val left right))
\end{verbatim}

Design a function {\tt all-words} that consumes a {\tt BSD} and
produces a list of all the words (just the words, not their meanings)
defined in the dictionary, in alphabetic order.

@@

(define a (make-defn "a" "first"))
(define b (make-defn "b" "second"))
(define c (make-defn "c" "third"))
(define l (make-leaf))

;; all-words : BSD -> [Listof String]
;; List all words defined in given dictionary in alphabetic order
(check-expect (all-words l) '())
(check-expect (all-words (make-node b (make-node a l l) (make-node c l l)))
              (list "a" "b" "c"))
(define (all-words bsd)
  (cond [(leaf? bsd) '()]
        [(node? bsd)
         (append (all-words (node-left bsd))
                 (list (defn-word (node-val bsd)))
                 (all-words (node-right bsd)))]))
\end{SaveVerbatim}

\noindent
{\bf Problem 3 (15 points).}
%
For each of the following functions, correctly reformulate the
function in terms of an existing abstraction function.  You may use
either lambda-expressions or local definitions if needed.

\begin{alltt}
;; any-zero? : [Listof Number] -> Boolean
;; Are any of the numbers in given list equal to zero?
(check-expect (any-zero? '(1 2 3)) #false)
(check-expect (any-zero? '(1 0 3)) #true)
(define (any-zero? xs)
  (cond [(empty? xs) #false]
        [(cons? xs)
         (or (zero? (first xs))
             (any-zero? (rest xs)))]))
\end{alltt}

\vspace{.8in}

\begin{SaveVerbatim}{VerbEnv}
SOLUTION:

(define (any-zero? xs) (ormap zero? xs))

\censor{%
\vspace{-.8in}
\noindent
\BUseVerbatim{VerbEnv}}

\begin{alltt}
(define DOT (circle 10 "solid" "black"))
(define MT (empty-scene 200 200))

;; draw-dots : [Listof Posn] -> Image
;; Draw dots for all given positions on an empty scene.
(check-expect (draw-dots '()) MT)
(check-expect (draw-dots (list (make-posn 10 20) (make-posn 20 30)))
              (place-image DOT 10 20 (place-image DOT 20 30 MT)))
(define (draw-dots ps)
  (cond [(empty? ps) MT]
        [(cons? ps)
         (place-image DOT
                      (posn-x (first ps))
                      (posn-y (first ps))
                      (draw-dots (rest ps)))]))
\end{alltt}

(define (draw-dots ps)
  (foldr (lambda (p img) (place-image DOT (posn-x p) (posn-y p) img))
         (empty-scene 200 200)
         ps))

@@

\newpage
\noindent
{\bf Problem 3 (cont).}

\begin{alltt}
;; remove-short-words : Number [Listof Defn] -> [Listof Defn]
;; Remove all definitions for words shorter than given length
(check-expect (remove-short-words 2 (list (make-defn "a" "first") 
                                          (make-defn "ab" "rest")))
              (list (make-defn "ab" "rest")))
(define (remove-short-words len ds)
  (cond [(empty? ds) '()]
        [(cons? ds)
         (if (< (string-length (defn-word (first ds))) len)
             (remove-short-words len (rest ds))
             (cons (first ds)
                   (remove-short-words len (rest ds))))]))
\end{alltt}

@@

(define (remove-short-words len ds)
  (filter (lambda (d) (not (< (string-length (defn-word d)) len))) ds))

@@

\noindent
{\bf Problem 4 (20 points).}
%
Design an abstraction of the following two functions and re-create
the original functions in terms of your abstraction.

\begin{verbatim}
;; sqr-pos : [Listof Number] -> [Listof Number]
;; Square each number in the list that is positive, remove the rest
(check-expect (sqr-pos '(2 -3 4)) '(4 16))
(define (sqr-pos xs)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (positive? (first xs))
             (cons (sqr (first xs))
                   (sqr-pos (rest xs)))
             (sqr-pos (rest xs)))]))

;; lowercase-lengths : [Listof String] -> [Listof Number]
;; Produce a list of lengths for all lowercase strings in given list
(check-expect (lowercase-lengths '("Hey" "you" "guys")) '(3 4))
(define (lowercase-lengths xs)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (string-lower-case? (first xs))
             (cons (string-length (first xs))
                   (lowercase-lengths (rest xs)))
             (lowercase-lengths (rest xs)))]))
\end{verbatim}

SOLUTION:
;; filter&map : [Listof X] [X -> Boolean] [X -> Y] -> [Listof Y]
(define (filter&map xs p f)
  (cond [(empty? xs) '()]
        [(cons? xs)
         (if (p (first xs))
             (cons (f (first xs))
                   (filter&map (rest xs) p f))
             (filter&map (rest xs) p f))]))

(define (sqr-pos xs)
  (filter&map xs positive? sqr))

(define (lowercase-lengths xs)
  (filter&map xs string-lower-case? string-length))
\end{SaveVerbatim}

\noindent
[Space for problem 4.]

\noindent
{\bf Problem 5 (20 points).}
%
Consider the following data definition (and examples) for representing sentences:
\begin{verbatim}
;; A Sentence is a [NEListof String]

(define sent1 (list "a" "dog" "barked"))
(define sent2 (list "a" "cat" "meowed"))
\end{verbatim}
The {\tt [NEListof X]} data definition is used to represent 
non-empty lists of elements:
\begin{verbatim}
;; A [NEListof X] is one of:
;; - (cons X '())
;; - (cons X [NEListof X])
\end{verbatim}
Given a {\tt Sentence}, a \emph{bigram} is a list of two consecutive
elements of the sentence. For example, {\tt sent1} has exactly two
bigrams:
\begin{verbatim}
(list "a" "dog")
(list "dog" "barked")
\end{verbatim}
Using the following data definition,
\begin{verbatim}
;; a Bigram is a (list String String)
\end{verbatim}
design a function {\tt bigrams} which takes a {\tt Sentence} and
returns a list of {\tt Bigrams}.

@@

;; bigrams : Sentence -> [Listof Bigram]
;; List of all adjacent words in a sentence
(check-expect (bigrams (list "one")) '())
(check-expect (bigrams sent1)
              (list (list "a" "dog") (list "dog" "barked")))
(define (bigrams sent)
  (cond [(empty? (rest sent)) '()]
        [else
         (cons (list (first sent) (second sent))
               (bigrams (rest sent)))]))
\end{SaveVerbatim}

@@

\newpage
\noindent
[Space for problem 5.]
